Maenpuen et al. Appendix S1-7: Sources and consequences of mismatch between leaf disc and whole-leaf leaf mass per area (LMA)

### 2022-04-08 ms/LMA_method.tex
add Bao's CV

\textbf{RESULTS:} Disc-based estimates of species and individual mean
LMA matched well whole-leaf estimates. Thin and large leaves had weaker
correlations between leaf disc and whole-leaf estimates. Variability of
LMA estimates within species (i.e., amount of intraspecific variation)
only weakly matched between whole-leaf and disc-based estimates
(\emph{R\textsuperscript{2}} = 0.31). We also found that sample with
small total dry mass of leaf discs tended to show greater differences
between leaf disc and whole-leaf LMA.

@@

Disc-based estimates of individual mean LMA also showed positive
correlations with whole-leaf estimates but their
\emph{R\textsuperscript{2}} values were smaller than those for the
species means (Fig. 2). Similar to the species-level pattern,
individuals with thicker and smaller leaves had stronger correlations
(Fig. 2). The relationships of the whole-leaf LMA : disc leaf LMA ratio
with total dry mass for the leaf disc (we weighted 2-3 discs together),
with leaf thickness, and with leaf area were heteroscedastic (unequal
variance, \emph{P} \textless{} 0.001; Appendix S3). Samples with small
total dry mass, thin leaves, and larger leaves tended to show greater
variance (Fig. 2; Appendix S3). The total dry mass of leaf disc was
positively correlated with both leaf thickness and leaf punch size
(Appendix S4). Amounts of intraspecific variation (CV; coefficient of
variation) only weakly matched between whole-leaf and disc-based
estimates (Fig. 3; \emph{R\textsuperscript{2}} = 0.31).

@@

Even though we used the same species, the same individuals and the same
trait (LMA) to estimate the extents of intraspecific trait variation
(ITV), the explained variance of the relationship between whole-leaf and
disc-based estimate was surprisingly small (Fig. 3;
\emph{R\textsuperscript{2}} = 0.31). This is probably because leaf disc
samples have more sources of ITV than whole-leaf samples (i.e.,
variation among leaf discs within the same leaves). Although there are
not many studies that quantified trait variation within a single leaf, a
previous study that used a giant leaved species \emph{Alocasia
macrorrhiza}, suggests that LMA varied within leaves due to reduced
water supply and demand from midrib to outer regions
(\protect\hyperlink{ref-Li2013a}{Li et al., 2013}). Because calculating
the mean values generally diminishes the effect of within-sample
variation, the relationship between disc-based estimates of individual
mean LMA and its whole-leaf estimates will be stronger than the
relationship between leaf disc CV and whole-leaf CV. Thus, leaf disc and
whole-leaf estimates of species means will show the strongest
relationship. This is consistent with the \emph{R\textsuperscript{2}}
values in our results (Figs. 1-3). The mismatch of ITV between
whole-leaf and disc-based estimate could be ignored in community-level
ITV studies (i.e., \protect\hyperlink{ref-Siefert2015}{Siefert et al.,
2015}), which quantify the total variance of ITV within communities.
However, it is likely to change patterns for species-level ITV studies
(i.e., \protect\hyperlink{ref-Albert2010a}{Albert et al., 2010}), which
quantify extents of ITV for each species, since species with high ITV
based on leaf-disc estimates do not always have to show high ITV for
whole-leaf estimates. This problem will be more obvious when one uses
smaller leaf punches (Appendix S7).

>>>>

\textbf{RESULTS:} Disc-based estimates of species and individual mean
LMA matched well whole-leaf estimates. Thin and large leaves had weaker
correlations between leaf disc and whole-leaf estimates. Variability of
LMA estimates within species (i.e., amount of intraspecific variation)
only weakly matched between whole-leaf and disc-based estimates
(\emph{R\textsuperscript{2}} = 0.32). We also found that sample with
small total dry mass of leaf discs tended to show greater differences
between leaf disc and whole-leaf LMA.

@@

Disc-based estimates of individual mean LMA also showed positive
correlations with whole-leaf estimates but their
\emph{R\textsuperscript{2}} values were smaller than those for the
species means (Fig. 2). Similar to the species-level pattern,
individuals with thicker and smaller leaves had stronger correlations
(Fig. 2). The relationships of the whole-leaf LMA : disc leaf LMA ratio
with total dry mass for the leaf disc (we weighted 2-3 discs together),
with leaf thickness, and with leaf area were heteroscedastic (unequal
variance, \emph{P} \textless{} 0.001; Appendix S3). Samples with small
total dry mass, thin leaves, and larger leaves tended to show greater
variance (Fig. 2; Appendix S3). The total dry mass of leaf disc was
positively correlated with both leaf thickness and leaf punch size
(Appendix S4). Amounts of intraspecific variation (CV; coefficient of
variation) only weakly matched between whole-leaf and disc-based
estimates (Fig. 3; \emph{R\textsuperscript{2}} = 0.32).

@@

Even though we used the same species, the same individuals and the same
trait (LMA) to estimate the extents of intraspecific trait variation
(ITV), the explained variance of the relationship between whole-leaf and
disc-based estimate was surprisingly small (Fig. 3;
\emph{R\textsuperscript{2}} = 0.32). This is probably because leaf disc
samples have more sources of ITV than whole-leaf samples (i.e.,
variation among leaf discs within the same leaves). Although there are
not many studies that quantified trait variation within a single leaf, a
previous study that used a giant leaved species \emph{Alocasia
macrorrhiza}, suggests that LMA varied within leaves due to reduced
water supply and demand from midrib to outer regions
(\protect\hyperlink{ref-Li2013a}{Li et al., 2013}). Because calculating
the mean values generally diminishes the effect of within-sample
variation, the relationship between disc-based estimates of individual
mean LMA and its whole-leaf estimates will be stronger than the
relationship between leaf disc CV and whole-leaf CV. Thus, leaf disc and
whole-leaf estimates of species means will show the strongest
relationship. This is consistent with the \emph{R\textsuperscript{2}}
values in our results (Figs. 1-3). The mismatch of ITV between
whole-leaf and disc-based estimate could be ignored in community-level
ITV studies (i.e., \protect\hyperlink{ref-Siefert2015}{Siefert et al.,
2015}), which quantify the total variance of ITV within communities.
However, it is likely to change patterns for species-level ITV studies
(i.e., \protect\hyperlink{ref-Albert2010a}{Albert et al., 2010}), which
quantify extents of ITV for each species, since species with high ITV
based on leaf-disc estimates do not always have to show high ITV for
whole-leaf estimates. This problem will be more obvious when one uses
smaller leaf punches (Appendix S7).
